---
title: "inegiR"
author: "Alan Ponce"
date: "12/2/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(inegiR)
library(knitr)

#Reference [http://enelmargen.org/inegiR/vignette_eng.html]
```


##  downloading a data series

```{r}

token <- "6b11996c-1a52-72a9-0187-e49bf9e607ce"

```

Now, I wish to find the rate of inflation (which in the case of INEGI is a percent change of the INPC data series).

This is the corresponding URL for INPC data.series:

```{r}

#XML format
urlINPC <- "http://www3.inegi.org.mx/sistemas/api/indicadores/v1//Indicador/216064/00000/es/false/xml/"

#JSON format
urlINPC2 <- "http://www3.inegi.org.mx/sistemas/api/indicadores/v1//Indicador/216064/00000/es/false/json/"

#Now, we are going to download this data series as a data.frame.

INPC <- serie_inegi(urlINPC, token)

# take a look
head(INPC)
tail(INPC)
glimpse(INPC)

```

The optional “metadata” parameter in serie_inegi allows us to also download the metadata information from the data series, which includes date of update, units, frequency, etc.

If “metadata” is set to TRUE, the information is parsed as a list of two elements: the metadata and the data frame.

```{r}
INPC_Metadata <- serie_inegi(urlINPC, token, metadata = TRUE)
#INPC_Metadata

class(INPC_Metadata)
glimpse(INPC_Metadata)

# To access any of these elements, simply use as a list:
# date of last update
INPC_Metadata$MetaData$UltimaActualizacion

```

Now that we have the INPC data series, we must apply a year-over-year change. For this we use the handy YoY() function, which let’s us choose the amount of periods to compare over (12 if you want year over year for a monthly series):

```{r}

Inflation <- YoY(INPC$Valores, 
                 lapso = 12, 
                 decimal=FALSE)

class(Inflation)
head(Inflation,  n= 20L)

# if we want a dataframe, we simply build like this
Inflation_df <- cbind.data.frame(Fechas = INPC$Fechas, 
                                 Inflation = Inflation)

Inflation_df
glimpse(Inflation_df)

```

This method works for any URL obtained from the INEGI documentation, but for the most used indicators, the package has built-in shortcut wrappers.

Let us obtain the same data series (inflation) via one of these specified shortcut functions:

```{r}
Inflation_fast <- inflacion_general(token)
#Inflation_fast

glimpse(Inflation_fast)
head(Inflation_fast)
tail(Inflation_fast)

```

## Downloading statistics from DENUE


```{r}

token_denue <- "31fae5dd-6bd4-4d33-a0e0-b0917cd55db2"

#Reference v2 [http://enelmargen.org/datascience/inegiR-v2/index.html]
```


To download the businesses in a certain radius, we need a few coordinates. Let’s use the ones around Monterrey Mexico’s main square:

```{r}
latitud_macro<-"25.669194"

longitud_macro<-"-100.309901"
```

Now, we download into a data.frame the list of businesses in a 250 meter radius.

```{r}

NegociosMacro <- denue_inegi(latitud = latitud_macro, 
                             longitud = longitud_macro, 
                             token_denue)

#NegociosMacro

head(NegociosMacro)[,1:2]


```

If you would like to change some parameters, this is accepted. For example a 1km radius and only businesses with “Restaurante” in the description.

```{r}

RestaurantsMacro <- denue_inegi(latitud = latitud_macro, 
                                longitud = longitud_macro, 
                                token_denue, 
                                metros = 1000, 
                                keyword = "Restaurante")


head(RestaurantsMacro)[,1:2]

```

```{r}

#token<-"webservice_token"
latitud<-"25.669194"
longitud<-"-100.30990"
# in english
businesses <- inegi_denue(latitud, longitud, token_denue, metros = 1000)
businesses
glimpse(businesses)
#View(businesses)
#kable(businesses)


# in spanish (legacy)
negocios <- denue_inegi(latitut, longitud, token_denue, metros = 100)


```

